Peptide retention time prediction

### 2014-11-23 retention.tex
Small fix in the Introduction

Today a majority of the mass spectrometry (MS) protocols in use are
high-throughput, generating extensive amounts of data. This has
brought up a whole new set of analytical challenges, making the
development of efficient algorithms to interpret this data a major
area of proteomics research. A particular difficult challenge in this
context is posed by the tremendous complexity of the data. Proteins
display a wide range of physico-chemical properties and are often
found at very different amounts in complex samples, spanning as much
as 12 orders of magnitude in relative abundance~\cite{Angel2012}.
Furthermore, a large proportion of proteins are subjected to
posttranslational modifications. As an example, a few years ago more
than 25,000 phosphorylation sites on human proteins were already
reported~\cite{Lemeer2009}. To manage this formidable diversity,
increasing efforts are made to extract as much information as possible
from an experiment.

The main purpose of liquid chromatography is to reduce the number of
peptides simultaneously injected into the mass spectrometer, and by
this increase the proteome coverage achieved in an experiment.
However, besides this, the separation process in itself provides
useful information about the analytes. On the one hand, the instrument
always records the time when a peptide elutes from the chromatography
column, i.e. the retention time. On the other hand, the retention time
can also be estimated by examining the sequence of the peptide. The
availability of both predicted and observed retention times has proven
to be a powerful feature, with several applications in mass
spectrometry-based proteomics \cite{Palmblad2013}.

>>>>

Today a majority of the mass spectrometry (MS) protocols in use are
high-throughput, generating extensive amounts of data. This has
brought up a whole new set of analytical challenges, making the
development of efficient algorithms to interpret this data a major
area of proteomics research. A particular difficult challenge in this
context is posed by the tremendous complexity of the data. Proteins
display a wide range of physico-chemical properties and are often
found at very different amounts in complex samples, spanning as much
as 12 orders of magnitude in relative abundance~\cite{Angel2012}.
Furthermore, a large proportion of proteins are subjected to
posttranslational modifications. As an example, a few years ago more
than 25,000 phosphorylation sites on human proteins were already
reported~\cite{Lemeer2009}. To manage this formidable diversity,
increasing efforts are made to extract - besides the fragmentation
spectra - additional information from each MS experiment.

An example of such additional information is represented by retention
times, i.e. the time points when peptides elute from the liquid
chromatography column.  The main purpose of liquid chromatography is
to reduce the number of peptides simultaneously injected into the mass
spectrometer, and by this increase the proteome coverage achieved in
an experiment.  However, besides this, the separation process in
itself provides useful information about the analytes. On the one
hand, the instrument always records the time when a peptide elutes
from the chromatography column, i.e. the retention time. On the other
hand, the retention time can also be estimated by examining the
sequence of the peptide. The availability of both predicted and
observed retention times has proven to be a powerful feature, with
several applications in mass spectrometry-based proteomics
\cite{Palmblad2013}.
